---
title: "Course Project 1 - Activity Monitoring Data"
author: "Rong Su"
date: "Friday, November 14, 2014"
output: html_document
---

First Step is to read the zip file from the course page and then unzip and read activity.csv file into R.
```{r echo=TRUE}
setwd("~/ReproducibleResearch")
fileURL <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileURL, destfile = "./activitymonitoring.zip",method = "auto")
dateDownloaded <- date()

if(!file.exists("data")) {
    dir.create("data")
}
zipdir <- "./data"
unzip("activitymonitoring.zip",exdir=zipdir)
files <- list.files(zipdir)
if(length(files)>1) stop("More than one data file inside.")
filepath <- paste(zipdir,files[1],sep="/")
activitydata <- read.csv(filepath)
```

The second step is to aggregate the data into steps per day and make a histogram of total number of steps taken each day.
```{r}
stepsperday <- aggregate(steps ~ date, data=activitydata, FUN=sum, na.action = na.omit)
hist(stepsperday$steps,main="Total Steps taken each day",
       xlab="Total Steps each day")
mean(stepsperday$steps)
median(stepsperday$steps)
```

Reshaping the data data by intervals, and steps are the average across all dates.
```{r}
stepsperinterval<- aggregate(steps ~ interval, data=activitydata, FUN=mean, na.action = na.omit)
```
We will now plot on 5 minute interval (x-axis) and average steps taken(across all days) as y-axis. As you may notice, there is a spike on the plot.
```{r}
plot(stepsperinterval,type="l")
```
As it turns out, the following 5 minute interval is the maximum spike.
```{r}
stepsperinterval[which.max(stepsperinterval$steps),1]
```

Now we are moving on to find out more about missing data. Total number of row with NAs in the original Activity Monitoring data are as follows.
```{r}
sum(!complete.cases(activitydata))
```
Now we will replace the NA value with imputed value - average steps take per interval. A new dataset with replaced values will be created here.
```{r}
meanstepsperinterval <- aggregate(steps ~ interval, data=activitydata, FUN=mean, na.action = na.omit)
activitynew <- activitydata
for(i in 1:dim(activitynew)[1])
  if(is.na(activitynew[i,1]) & length(meanstepsperinterval[which(meanstepsperinterval$interval==activitynew[i,3]),2])>0)
      activitynew[i,1] <- meanstepsperinterval[which(meanstepsperinterval$interval==activitynew[i,3]),2]
```
We will re-calculate the total number of steps per day, mean, and median based on the new data after replacing NAs.

```{r}
newstepsperday <- aggregate(steps ~ date, data=activitynew, FUN=sum)
hist(newstepsperday$steps,main="Imputed total Steps taken each day",
       xlab="Imputed Total Steps each day")
mean(newstepsperday$steps)
median(newstepsperday$steps)
```
Now we are adding a new factor colunm to the new dataset to identify whether it is weekday or weekend.
```{r}
activitynew$weekdays <- ifelse(weekdays(as.Date(activitynew$date)) != "Sunday" & weekdays(as.Date(activitynew$date)) != "Saturday", "weekday","weekend")
newstepsperinterval<- aggregate(steps ~ interval + weekdays, data=activitynew, FUN=mean)

library(lattice)
xyplot(steps~interval|weekdays, group=weekdays, data=newstepsperinterval,type='l', layout=c(1,2))

```